### 2011-05-01 talks/talk.tex

\documentclass[utf8,handout]{beamer}

\usepackage{graphicx}
\usepackage{color}
\usepackage{listings}
\usetheme{Boadilla}

\setbeamertemplate{footline}%{infolines theme}
{
\leavevmode%
\hbox{%
\begin{beamercolorbox}[wd=.333333\paperwidth,ht=2.25ex,dp=1ex,center]{author in head/foot}%
\usebeamerfont{author in head/foot}\insertshortauthor%~~(\insertshortinstitute)
\end{beamercolorbox}%
\begin{beamercolorbox}[wd=.333333\paperwidth,ht=2.25ex,dp=1ex,center]{title in head/foot}%
\usebeamerfont{title in head/foot}\insertshorttitle
\end{beamercolorbox}%
\begin{beamercolorbox}[wd=.333333\paperwidth,ht=2.25ex,dp=1ex,right]{date in head/foot}%
\usebeamerfont{date in head/foot}\insertshortdate{}\hspace*{2em}
\insertframenumber{} / \inserttotalframenumber\hspace*{2ex}
\end{beamercolorbox}}%
\vskip0pt%
}

\begin{document}

\title[KiCS2]{KiCS2: A New Compiler from Curry to Haskell}

\date{May 04, 2011}

\author[B. Braßel, M. Hanus, B. Peemöller, F. Reck]
 {Bernd Braßel \and Michael Hanus \\
   Björn Peemöller \and Fabian Reck\\
  \texttt{\{bbr, mh, bjp, fre\}@informatik.uni-kiel.de}}

\institute{Kiel University}

\begin{frame}
  \titlepage
\end{frame}

\section{Introduction}

\begin{frame}[fragile]
\frametitle{Curry}
\begin{columns}
\column{.45\textwidth}
\begin{itemize}
\item Lazy functional logic programing language.
\item Haskell-like syntax.
\item Extended by non-determinism and free variables. %TODO: failure?
\end{itemize}
\column{.45\textwidth}
\begin{block}{Running example}
\begin{semiverbatim}
data Bool = True | False

aBool :: Bool
aBool = True ? False

not True  = False
not False = True

xor True  x = not x
xor False x = x

xorSelf x = xor x x
\end{semiverbatim}
\end{block}
\end{columns}
\end{frame}

\begin{frame}[fragile]
\frametitle{Example Call}
\begin{semiverbatim}
> xorSelf aBool
False
False
No more solutions.
\end{semiverbatim}
\begin{itemize}
\item The call of \verb!aBool! yields two solutions
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{KiCS2}
% Compilation strategies
%   - compile to logic language (Pakcs compiles to Prolog)
%   - compile to functional language (KiCS compiles to Haskell)
%   - compile to abstract machine (MCC compiles to C)
\begin{block}{Approach}
\begin{itemize}
\item Compile Curry programs into Haskell
\item Reuse expresiveness of Haskell (lazy evaluation, higher-order functions)
\item Benefit from mature Haskell compiler (GHC)
\end{itemize}
\end{block}

\begin{block}{Goals}
\begin{itemize}
\item Efficient execution of purely functional programs
\item Avoid unsafe features to enable optimization (in contrast to KiCS)
\item Support for different search strategies (including parallel strategies)
\end{itemize}
\end{block}
\end{frame}

\section{Compilation scheme}

\begin{frame}[fragile]
\frametitle{Representation of Non-Determinism}
\begin{itemize}
  \item No built-in non-determinism in Haskell.
  \item Idea: Explicit representation in data types as constructors.
\end{itemize}

\begin{block}{Curry}
\begin{semiverbatim}
data Bool = True | False

aBool :: Bool
aBool = True ? False
\end{semiverbatim}
\end{block}

\begin{block}{Haskell}
\begin{semiverbatim}
data Bool = True | False | Choice_Bool Bool Bool

aBool :: Bool
aBool = Choice_Bool True False
\end{semiverbatim}
\end{block}
\end{frame}

\begin{frame}[fragile]
\frametitle{Extension of the Operations}
\begin{itemize}
  \item Operations have to deal with non-deterministic values.
  \item Non-determinism in arguments is propagated.
%   \item Purely functional Curry programs are Haskell programs (no runtime overhead) (first-order)
\end{itemize}

\begin{block}{Haskell}
\begin{semiverbatim}
not True                = False
not False               = True
not (Choice_Bool x1 x2) = Choice_Bool (not x1) (not x2)

xor True                x = not x
xor False               x = x
xor (Choice_Bool x1 x2) x = Choice_Bool (xor x1 x)
                                        (xor x2 x)

xorSelf x = xor x x
\end{semiverbatim}
\end{block}
\end{frame}

\begin{frame}[fragile]
\frametitle{Example calls}
\begin{example}
\verb!not aBool!
$\leadsto$ \verb!not (Choice_Bool True False)!
$\leadsto$ \verb!Choice_Bool (not True) (not False)!
$\leadsto$ \verb!Choice_Bool False True!
\end{example}

\begin{example}
\verb!xorSelf aBool!
$\leadsto$ \verb!xor aBool aBool!
$\leadsto$ \verb!xor (Choice_Bool True False) aBool!
$\leadsto$ \verb!Choice_Bool (xor True aBool) (xor False aBool)!
$\leadsto$ \verb!Choice_Bool (not aBool) aBool!
$\leadsto$ \verb!Choice_Bool (Choice_Bool False True) (Choice_Bool True False)!
\end{example}

\begin{itemize}
  \item Yiels four results: \verb!False, True, True, False!
  \item Only \verb!False, False! are reasonable.
\end{itemize}

\end{frame}

\begin{frame}[fragile]
\frametitle{Run-Time Choice vs. Call-Time Choice}
\begin{block}{Curry}
\begin{lstlisting}
xorSelf x = xor x x
\end{lstlisting}
\end{block}
\pause
\begin{block}{Run-Time Choice}
\begin{itemize}
\item Variables denote non-deterministic computations
\item {\tt x}'s may evaluate to different values in the same non-deterministic branch
\end{itemize}
\end{block}
\pause
\begin{block}{Call-Time Choice}
\begin{itemize}
\item Variables denote values
\item Both {\tt x} evaluate to the same value in each non-deterministic branch
\end{itemize}
\end{block}
\pause
\begin{itemize}
\item Call-Time Choice is often more intuitive
\item Curry has Call-Time Choice semantics 
\item Unrestricted propagation violates Call-Time Choice
\end{itemize}
\end{frame}

\begin{frame}[fragile]
\frametitle{Identifying Choices}
\begin{itemize}
\item Choices are annotated with unique identifiers
\item Functions that may introduce non-determinism are supplied with an additional parameter to obtain fresh \verb!ID!s
\item Decisions made for different occurrences of the same Choice have to be consistent
\end{itemize}

\begin{block}{Choice Identifier}
\begin{semiverbatim}
type ID = Integer

data Bool = False | True | Choice_Bool ID Bool Bool

aBool :: ID -> Bool
aBool i = Choice_Bool i True False

xor (Choice_Bool i x1 x2) x = Choice_Bool i (xor x1 x) 
                                            (xor x2 x)
\end{semiverbatim}
\end{block}
\end{frame}

\begin{frame}[fragile]
\frametitle{Example call}
% Mention that (aBool 1) is computed only once due to sharing
% Mention the origin of the initial ID 1
\begin{example}
\verb!xorSelf (aBool 1)!
$\leadsto$ \verb!xor (aBool 1) (aBool 1)!
$\leadsto$ \verb!xor (Choice_Bool 1 True False) (Choice_Bool 1 True False)!
$\leadsto$ $\dots$
$\leadsto$ \verb!Choice_Bool 1 (Choice_Bool 1 False True) (Choice_Bool 1 True False)!
\end{example}

\begin{itemize}
\item The results \verb!True, True! are identified as invalid
\end{itemize}
\end{frame}

\begin{frame}[fragile]
\frametitle{Representing Failure}
\begin{itemize}
 \item Computations with failures is a common programming pattern in FLP
 \item Failures do not abort the computation, but are ignored
 \item Thus, failures have to be represented as well.
\end{itemize}

\begin{block}{Curry}
\begin{semiverbatim}
data Bool = True | False

ensureTrue True = True
\end{semiverbatim}
\end{block}

\begin{block}{Haskell}
\begin{semiverbatim}
data Bool = True | False | Choice_Bool Bool Bool | Fail_Bool

ensureTrue True                  = True
ensureTrue (Choice_Bool i x1 x2) = ...
ensureTrue _                     = Fail_Bool
\end{semiverbatim}
\end{block}
\end{frame}

\begin{frame}[fragile]
\frametitle{Higher Order Functions}
\begin{itemize}
\item Up to now only first-order programs were considered
\item Deterministic higher-order functions can remain unchanged
\end{itemize}
\begin{columns}
\column{.45\textwidth}
\begin{block}{Curry}
\begin{semiverbatim}
apply :: (a -> b) -> a -> b
apply f x = f x

> apply not True
False
\end{semiverbatim}
\end{block}
\column{.45\textwidth}
\begin{block}{Haskell}
\begin{semiverbatim}
apply :: (a -> b) -> a -> b
apply f x = f x

> apply not True
False
\end{semiverbatim}
\end{block}
\end{columns}
\end{frame}

\begin{frame}[fragile]
\frametitle{Higher Order Functions}
\begin{itemize}
\item Non-deterministic functions need an additional parameter
\item An identifier is passed to functional argument when applied
\end{itemize}
\begin{columns}
\column{.45\textwidth}
\begin{block}{Curry}
\begin{semiverbatim}
apply :: (a -> b) -> a -> b
apply f x = f x

ndNot x = x ? not x

> apply ndNot True
True
False
\end{semiverbatim}
\end{block}
\column{.45\textwidth}
\begin{block}{Haskell}
\begin{semiverbatim}
apply :: (a -> ID -> b) 
      -> a -> ID -> b
apply f x i = f x i

ndNot x i = 
  Choice_Bool i x (not x)

> apply ndNot True 1
Choice_Bool 1 True False
\end{semiverbatim}
\end{block}
\end{columns}
\end{frame}

\begin{frame}[fragile]
\frametitle{Higher-order optimization}
Deterministic functions have to be wrapped
\begin{block}{Curry}
\begin{semiverbatim}
> apply not True
False
\end{semiverbatim}
\end{block}
\begin{block}{Haskell}
\begin{semiverbatim}
> apply (\\x _ -> not x) True 1
False
\end{semiverbatim}
\end{block}
\begin{block}{Optimization}
\begin{itemize}
\item Higher-order functions are translated into
  \begin{itemize}
   \item a deterministic variant
   \item a non-deterministic variant
  \end{itemize}
\item Benchmarks show enormously improved performance
\end{itemize}
\end{block}
Non-deterministic choice between functions is also possible.
\end{frame}

\section{Benchmarks}

\begin{frame}
\frametitle{Benchmarks}
\end{frame}

\section{Conclusion}

\begin{frame}
\frametitle{Additional Features}
\begin{itemize}

\item Free Variables
\item Unification
\item Function Patterns
\item Encapsulated Search

\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Conclusion}
\begin{itemize}
\item A transformation of a functional logic program into a functional program.
\item The resulting program can be optimized by the Haskell compiler
\item A formal description of our approach can be found in our paper.
\end{itemize}
\end{frame}

\begin{frame}
\begin{center}

 {\huge Thank You!}\\[3ex]
 \pause
 {\huge Questions?}

\end{center}
\end{frame}

\end{document}