### 2011-07-06 papers/wlp2011/paper.tex
benchmark update

\documentclass{llncs}

\addtolength{\textheight}{0.2cm}
\addtolength{\topmargin}{-0.1cm}
\addtolength{\textwidth}{0.2cm}
\addtolength{\oddsidemargin}{-0.1cm}
\addtolength{\evensidemargin}{-0.1cm}

\usepackage{times}
\usepackage{url}
\def\UrlFont{\tt}
\usepackage{latexsym}
\usepackage{xspace}
\usepackage{hyperref}
\usepackage{color}

\usepackage{listings}
\lstset{aboveskip=0.5ex,
        belowskip=0.5ex,
        showstringspaces=false, % no special string space
        mathescape=true,
        flexiblecolumns=false,
        basewidth=0.52em,
        basicstyle=\small\ttfamily}

\lstset{literate={->}{{$\rightarrow{}\!\!\!$}}3
       }

\lstnewenvironment{curry}{\lstset{backgroundcolor=\color[rgb]{0.9,0.9,0.9}}}{}
\lstnewenvironment{haskell}{}{}
\newcommand{\listline}{\vrule width0pt depth1.75ex}

\newcommand{\code}[1]{\mbox{\small\texttt{#1}}}
\newcommand{\ccode}[1]{``\code{#1}''}
\newcommand{\bs}{\char92\xspace} % backslash
\newcommand{\us}{\char95\xspace} % underscore
\newcommand{\dollar}{\code{\char36}} % dollar in programs
\newcommand{\Choice}[2]{#1 \mathop{?} #2}
\newcommand{\todo}[1]{\fbox{\sc To do: #1}}

% Redefine description env as equation list to gain key alignment
\usepackage{eqlist}
\let\description=\eqlist
\let\enddescription=\endeqlist
\let\eqlistlabel\descriptionlabel

\begin{document}
\pagestyle{plain}
\sloppy

\title{Implementing Equational Constraints\\ in a Functional Language}

\author{
Bernd Bra{\ss}el
\kern1em
Michael Hanus
\kern1em
Bj{\"o}rn Peem{\"o}ller
\kern1em
Fabian Reck
}
\institute{
Institut f\"ur Informatik, CAU Kiel, D-24098 Kiel, Germany \\
\email{\{bbr|mh|bjp|fre\}@informatik.uni-kiel.de}
}

\maketitle

\begin{abstract}
KiCS2 is a new system
to compile functional logic programs of the source language Curry
into purely functional Haskell programs.
The implementation is based on the idea to represent
the search space as a data structure and logic variables
as operations that generate their values.
This has the advantage that one can apply various,
and in particular, complete search strategies to compute solutions.
However, the generation of all values for logic variables
might be inefficient for applications that
exploit constraints on partially known values.
To overcome this drawback, we propose new techniques
to implement equational constraints in this framework.
In particular, we show how unification modulo function evaluation
and functional patterns can be added without sacrificing
the efficiency of the kernel implementation.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:Introduction}

Functional logic languages combine the most important
features of functional and logic programming in a single language
(see \cite{AntoyHanus10CACM,Hanus07ICLP} for recent surveys).
In particular, they provide higher-order functions and demand-driven
evaluation from functional programming together with logic programming features
like non-deterministic search and computing with partial information
(logic variables).
This combination
has led to new design patterns \cite{AntoyHanus02FLOPS,AntoyHanus11WFLP}
and better abstractions for application programming, but it also gave rise
to new implementation challenges.

Previous implementations of functional logic languages
can be classified into three categories:
\begin{enumerate}
 \item designing new abstract machines appropriately supporting
these operational features and implementing them in some (typically,
imperative) language, like C \cite{Lux99FLOPS}
or Java \cite{AntoyHanusLiuTolmach05,HanusSadre99JFLP},

 \item compilation into logic languages (Prolog) and reusing
the existing backtracking implementation for non-deterministic
search as well as logic variables and unification for computing with partial
information \cite{AntoyHanus00FROCOS,Lopez-FraguasSanchez-Hernandez99}, or

 \item compilation into non-strict functional languages like Haskell
and reusing the implementation
of lazy evaluation and higher-order functions
\cite{BrasselHuch07,BrasselHuch09}.
\end{enumerate}
%
The latter approach requires the implementation
of non-deterministic computations in a deterministic language
but has the advantage
that the explicit handling of non-determinism allows for
various search strategies like depth-first, breadth-first, parallel,
or iterative deepening instead of committing to a fixed (incomplete)
strategy like backtracking \cite{BrasselHuch07}.

In this paper we consider KiCS2 \cite{BrasselFischerHanusReck11},
a new system that compiles functional logic programs of the
source language Curry \cite{Hanus06Curry}
into purely functional Haskell programs.
We have shown in \cite{BrasselFischerHanusReck11}
that this implementation can compete with or outperform
other existing implementations of Curry.
KiCS2 is based on the idea to represent
the search space, i.e., all non-deterministic results of a computation,
as a data structure that can be traversed by operations
implementing various strategies.
Furthermore, logic variables are replaced by generators,
i.e., operations that non-deterministically evaluate to
all possible ground values of the type of the logic variable.
It has been shown \cite{AntoyHanus06ICLP}
that computing with logic variables by narrowing \cite{Reddy85,Slagle74}
or computing with generators by rewriting
are equivalent, i.e., compute the same values.
Although this implementation technique is correct \cite{Brassel11Thesis},
the generation of all values for logic variables
might be inefficient for applications that
exploit constraints on partial values.
For instance, in Prolog the equality constraint \ccode{X=c(a)}
is solved by instantiating the variable \code{X} to \code{c(a)},
but the equality constraint \ccode{X=Y} is solved by binding \code{X}
to \code{Y} without enumerating any values for \code{X} or \code{Y}.
Therefore, we propose in this paper new techniques
to implement equational constraints in the framework of KiCS2
(note that, in contrast to Prolog, unification is performed
modulo function evaluation).
Furthermore, we also show how functional patterns \cite{AntoyHanus05LOPSTR},
i.e., patterns containing evaluable operations providing
for more powerful pattern matching than in logic or functional languages,
can be implemented in this framework.
We show that both extensions lead to efficiency improvements
without sacrificing the efficiency of the kernel implementation.

In the next section, we review the source language Curry
and the features that are considered in this paper.
Section~\ref{sec:Compilation} sketches the implementation scheme
of KiCS2.
Sections~\ref{sec:Unification} and~\ref{sec:FuncPatterns}
discuss the extensions to implement unification modulo
functional evaluation and functional patterns, respectively.
Benchmarks demonstrating the usefulness of this scheme
are presented in Section~\ref{sec:Benchmarks}
before we conclude in Section~\ref{sec:Conclusions}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Curry Programs}
\label{sec:Curry}

The syntax of the functional logic language Curry \cite{Hanus06Curry}
is close to Haskell \cite{PeytonJones03Haskell},
i.e., type variables and names of defined operations usually
start with lowercase letters and the names of type and data constructors
start with an uppercase letter. The application of $f$
to $e$ is denoted by juxtaposition (``$f~e$'').
In addition to Haskell, Curry allows free (logic)
variables in conditions and right-hand sides of defining rules.
Hence, an operation is defined by conditional rewrite rules of the form:
%
\begin{equation}
\label{rule}
f~t_1 \ldots t_n \code{~|~} c \code{~=~} e \code{~~where~} vs \code{~free}
\end{equation}
where the \emph{condition} $c$ is optional and
$vs$ is the list of variables occurring in $c$ or $e$ but not in the
\emph{left-hand side} $f~t_1 \ldots t_n$.

In contrast to functional programming and similarly to logic programming,
operations can be defined by overlapping rules so that
they might yield more than one result on the same input.
Such operations are also called \emph{non-deterministic}.
For instance, Curry offers a \emph{choice} operation that is predefined by
the following rules:
%
\begin{curry}
  x ? _ = x
  _ ? y = y
\end{curry}
%
Thus, we can define a non-deterministic operation \code{aBool} by
\label{ex:aBool}
%
\begin{curry}
  aBool = True ? False
\end{curry}
%
so that the expression \ccode{aBool} has two values:
\code{True} and \code{False}.

If non-deterministic operations are used as arguments in other operations,
a semantical ambiguity might occur. Consider the operations
%
\begin{curry}
  xor True  False = True
  xor True  True  = False
  xor False x     = x

  xorSelf x = xor x x
\end{curry}
%
and the expression \ccode{xorSelf aBool}.
If we interpret this program as a term rewriting system,
we could have the reduction
\begin{haskell}
  xorSelf aBool  $\to~$  xor aBool aBool     $\to~$  xor True aBool
                 $\to~$  xor True False      $\to~$  True
\end{haskell}
leading to the unintended result \code{True}.
Note that this result cannot be obtained if we use a strict strategy
where arguments are evaluated prior to the function calls.
In order to avoid dependencies on the evaluation strategies
and exclude such unintended results,
Gonz\'alez-Moreno et al.\ \cite{GonzalezEtAl99} proposed
the rewriting logic CRWL as a logical
(execution- and strategy-independent) foundation for declarative
programming with non-strict and non-deterministic operations.  This
logic specifies the \emph{call-time choice} semantics \cite{Hussmann92}
\label{ctc-semantics}
where values of the arguments of an operation are determined before the
operation is evaluated. This can be enforced in a lazy strategy
by sharing actual arguments.
For instance, the expression above can be lazily evaluated
provided that all occurrences of \code{aBool}
are shared so that all of them reduce either to \code{True} or to \code{False}
consistently.

The condition $c$ in rule (\ref{rule}) is typically a
conjunction of \emph{equational constraints}
of the form \code{$e_1$\,=:=\,$e_2$}.
Such a constraint is satisfiable if
both sides $e_1$ and $e_2$ are reducible to unifiable data terms.
For instance, if the symbol ``\code{++}'' denotes the usual list
concatenation operation, we can define an operation \code{last}
that computes the last element \code{e} of a non-empty list \code{xs}
as follows:
%
\label{ex:last}
\begin{curry}
  last xs | ys++[e] =:= xs  = e   where ys, e free
\end{curry}
%
Like in Haskell, most rules defining functions are \emph{constructor-based} 
\cite{ODonnell85}, i.e., in (\ref{rule})
$t_1,\ldots,t_n$ consist of variables and/or data constructor symbols
only.  However, Curry also allows \emph{functional patterns}
\cite{AntoyHanus05LOPSTR}, i.e., $t_i$ might additionally contain calls to
defined operations. For instance, we can define the last element
of a list also by:
%
\begin{curry}
  last' (xs++[e]) = e
\end{curry}
%
Here, the functional pattern \code{(xs++[e])}
states that \code{last'} is reducible
to \code{e} provided that the argument can be matched against
some value of \code{(xs++[e])} where \code{xs} and \code{e} are
free variables.
By instantiating \code{xs} to arbitrary lists, the value
of \code{(xs++[e])} is any list having \code{e} as its last element.
Functional patterns are
a powerful feature to express arbitrary selections in term structures.
For instance, functional patterns support a straightforward
processing of XML data with incompletely specified
or evolving formats \cite{Hanus11ICLP}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The Compilation Scheme of KiCS2}
\label{sec:Compilation}

To understand the extensions described in the subsequent sections,
we sketch the translation of Curry programs
into Haskell programs as performed by KiCS2.
More details about this translation scheme can be found in
\cite{BrasselFischer08IFL,BrasselFischerHanusReck11}.

As mentioned in the introduction, the KiCS2 implementation
is based on the explicit representation of non-deterministic results
in a data structure.
This is achieved by extending each data type of the source program
by constructors to represent a choice between two values
and a failure, respectively.
For instance, the data type for Boolean values defined in a Curry program by
\begin{curry}
  data Bool = False | True
\end{curry}
is translated into the Haskell data type\footnote{Actually,
our compiler performs some renamings to avoid conflicts with
predefined Haskell entities and introduces type classes
to resolve overloaded symbols like \code{Choice} and \code{Fail}.}
\begin{haskell}
  data Bool = False | True | Choice ID Bool Bool | Fail
\end{haskell}
The first argument of type \code{ID} of each \code{Choice} constructor
is used to implement the call-time choice semantics
discussed in Section~\ref{sec:Curry}.
Since the evaluation of \code{xorSelf aBool} duplicates
the argument operation \code{aBool}, we have to ensure
that both duplicates, which later evaluate to a non-deterministic
choice between two values, yield either \code{True} or \code{False}.
This is obtained by assigning a unique identifier (of type \code{ID})
to each \code{Choice}. The difficulty is to get a unique identifier
when needed, i.e., when some operation evaluates to a \code{Choice}.
Since we want to compile into \emph{purely} functional programs
(in order to enable powerful program optimizations),
we cannot use unsafe features with side effects to generate
such identifiers.
Hence, we pass a (conceptually infinite) set of identifiers,
also called \emph{identifier supply},
to each operation so that a \code{Choice} can pick an identifier from
this set.
For this purpose, we assume a type \code{IDSupply},
representing an infinite set of identifiers,
with operations
\begin{haskell}
  initSupply  :: IO IDSupply
  thisID      :: IDSupply -> ID
  leftSupply  :: IDSupply -> IDSupply
  rightSupply :: IDSupply -> IDSupply
\end{haskell}
The operation \code{initSupply} creates such a set (at the beginning
of an execution),
the operation \code{thisID} takes some identifier from this set, and
\code{leftSupply} and \code{rightSupply} split this set
into two disjoint subsets without the identifier
obtained by \code{thisID}.
There are different implementations available \cite{AugustssonRittriSynek94}
(see below for a simple implementation) and our implementation
is parametric over concrete implementations of \code{IDSupply}.

When translating Curry to Haskell, KiCS2 adds to each operation
an additional argument of type \code{IDSupply}.
For instance, the operation \code{aBool}
defined in Section~\ref{ex:aBool} is translated into:
\begin{haskell}
  aBool :: IDSupply -> Bool
  aBool s = Choice (thisID s) True False
\end{haskell}
Similarly, the operation
\begin{curry}
  main :: Bool
  main = xorSelf aBool
\end{curry}
is translated into
\begin{haskell}
  main :: IDSupply -> Bool
  main s = xorSelf (aBool (leftSupply s)) (rightSupply s)
\end{haskell}
so that the set \code{s} is split into a set \code{(leftSupply s)}
containing identifiers for the evaluation of \code{aBool}
and a set \code{(rightSupply s)} containing identifiers
for the evaluation of the operation \code{xorSelf}.

Since all data types are extended by additional constructors,
we must also extend the definition of operations performing
pattern matching. For instance, consider the definition of polymorphic lists
\begin{curry}
  data List a = Nil | Cons a (List a)
\end{curry}
and an operation to extract the first element of a non-empty list:
\begin{curry}
  head :: List a -> a
  head (Cons x xs) = x
\end{curry}
The type definition is then extended as follows:
\begin{haskell}
  data List a = Nil | Cons a (List a) | Choice ID (List a) (List a) | Fail
\end{haskell}
The operation \code{head} is extended by an identifier supply
and further matching rules:
\begin{haskell}
  head :: List a -> IDSupply -> a
  head (Cons x xs)      s = x
  head (Choice i x1 x2) s = Choice i (head x1 s) (head x2 s)
  head _                s = Fail
\end{haskell}
The second rule transforms a non-deterministic argument
into a non-deterministic result and
the final rule returns \code{Fail} in all other cases,
i.e., if \code{head} is
applied to the empty list as well as if the matching argument
is already a failed computation (failure propagation).

To show a concrete example, we use the following implementation
of \code{IDSupply} based on unbounded integers:
%
\begin{haskell}
  type IDSupply = Integer
  initSupply    = return 1
  thisID      n = n
  leftSupply  n = 2 * n
  rightSupply n = 2 * n + 1
\end{haskell}
%
If we apply the same transformation to the rules defining \code{xor}
(actually, \code{xor} performs pattern matching on two arguments
which is avoided in KiCS2 by introducing an intermediate operation,
see \cite{BrasselFischerHanusReck11}) and evaluate the main expression
\code{(main 1)},
we obtain the result
\begin{haskell}
  Choice 2 (Choice 2 False True) (Choice 2 True False)
\end{haskell}
Thus, the result is non-deterministic and contains several choices,
but all of them have the same identifier.
This identifier is used when the values contained in this result
are searched for: one has to make \emph{consistent} selections in choices with
same identifiers. Thus, if we select the left branch as the value
of the outermost \code{Choice}, we also have to select the left branch
in the selected argument \code{(Choice 2 False True)} so that only
the value \code{False} is possible here.
Similarly, if we select the right branch as the value of the outermost
\code{Choice}, we also have to select the right branch in
its selected argument \code{(Choice 2 True False)} which yields the only
value \code{False}.
Thus, the unintended value \code{True} is not produced.

In general, to extract all values from a \code{Choice} structure,
we have to traverse it, compute all possible choices
but consider the choice identifiers to make consistent (left/right)
decisions.
The latter requirement can be implemented by storing
the decisions already made for some choices during the traversal.
For this purpose, we introduce the type
\begin{haskell}
  data Decision = NoDecision | ChooseLeft | ChooseRight
\end{haskell}
where \code{NoDecision} represents the fact that the value of a choice
is yet undecided.
Furthermore, we assume operations to lookup the current decision
for a given identifier or change its decision (depending on the implementation
of \code{IDSupply}, KiCS2 supports various implementations
based on memory cells or finite maps):
\begin{haskell}
  lookupDecision :: ID -> IO Decision
  setDecision    :: ID -> Decision -> IO ()
\end{haskell}
%
Now we can print all values contained in a choice structure
in a depth-first manner by the following I/O operation:\footnote{%
Note that this code has been simplified for readability
since the type system of Haskell does not
allow this direct definition.}
\label{sec:printValsDFS}
\begin{haskell}
  printValsDFS :: a -> IO ()
  printValsDFS Fail             = return ()
  printValsDFS (Choice i x1 x2) = lookupDecision i >>= follow
  printValsDFS v                = print v
   where
    follow ChooseLeft  = printValsDFS x1
    follow ChooseRight = printValsDFS x2
    follow NoDecision  = do newDecision ChooseLeft  x1
                            newDecision ChooseRight x2

    newDecision d x = do setDecision i d
                         printValsDFS x
                         setDecision i NoDecision
\end{haskell}
This operation ignores failures and prints values that are not a \code{Choice}.
For a \code{Choice}, it checks whether a decision for
this identifier has already been made (note that the initial value
for all identifiers is \code{NoDecision}).
If a decision has been made for this choice, it follows this decision.
Otherwise, the left alternative is used and this decision is stored.
After printing
all the values w.r.t.\ this decision,
the decision is undone (like in backtracking)
and the right alternative is used and stored.

In general, this operation is applied to the normal form
of the main expression (where \code{initSupply} is used to
compute an initial identifier supply passed to this expression).
The normal form computation is necessary for structured data like lists
so that a failure or choice in some part of the data is moved to
the root. Other search strategies, like
breadth-first search, iterative deepening, or parallel search,
can be obtained by different implementations of this main operation
to print all values. Instead of printing all values,
one can also collect all values in a tree-like data structure
so that the programmer can implement his own search strategies
(this corresponds to encapsulating search \cite{BrasselHanusHuch04JFLP}).
Finally, one can also have other options rather than printing all values
of the main expression.
For instance, one can easily define operations
to print either the first solution only or one by one upon user request.
Due to the lazy evaluation strategy of Haskell,
such operations can also be applied to infinite choice structures.

To avoid an unnecessary growth of the search space represented by
\code{Choice} constructors, our compiler performs an optimization for
deterministic operations. If an operation does not call, 
neither directly nor indirectly through
other operations, the choice operation \ccode{?},\footnote{%
Note that before applying this compilation scheme, overlapping rules in
user-defined operations are transformed into calls of the choice operation
\ccode{?}.}
then it is not necessary to pass an identifier supply to the operation.
In consequence, only the matching rules are extended so that the generated code
is nearly identical to a corresponding functional program.
Actually, the benchmarks presented in \cite{BrasselFischerHanusReck11}
show that this implementation outperforms all other Curry implementations
for deterministic operations,
and, for non-deterministic operations, outperforms Prolog-based
implementations of Curry and can compete with MCC \cite{Lux99FLOPS},
a Curry implementation that compiles to C.

As mentioned in the introduction,
KiCS2 translates occurrences of logic variables
into generators. For instance, consider the operation \code{not}
defined by the rules:
\begin{curry}
  not True  = False
  not False = True
\end{curry}
Then the expression \ccode{not x}, where \code{x} is a logic variable,
is translated into \ccode{not aBool} (we omit the \code{IDSupply} argument
here). The latter expression is evaluated by reducing the
argument \code{aBool} to a choice between \code{True} or \code{False}
followed by applying \code{not} to this choice.
This is similar to a narrowing step on \ccode{not x}
that instantiates the variable \code{x} to \code{True} or \code{False}.
% Actually, it has been shown in \cite{AntoyHanus06ICLP}
% that narrowing expressions with logic variables
% is equivalent to rewriting with generators.
Since such generators are standard
non-deterministic operations, they are translated like any other operation
so that one does not need additional run-time support for them.
However, if equational constraints should be solved,
there are methods which are more efficient than generating all values.
These methods and their implementation are discussed in the
next section.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Equational Constraints and Unification}
\label{sec:Unification}

As known from logic programming, predicates or constraints
are important to restrict the set of intended values
in a non-deterministic computation.
Apart from user-defined predicates,
equational constraints of the form \code{$e_1$\,=:=\,$e_2$}
are the most important kind of constraints.
We have already seen a typical application
of an equational constraint in the operation
\code{last} in Section~\ref{ex:last}.

Due to the presence of non-terminating operations
and infinite data structures,
\ccode{=:=} is interpreted as the \emph{strict equality} on terms
\cite{GiovannettiLeviMoisoPalamidessi91},
i.e., the equation \code{$e_1$\,=:=\,$e_2$} is satisfied iff
$e_1$ and $e_2$ are reducible to the same ground constructor term.
In particular, expressions that do not have a value
are not equal w.r.t.\ \ccode{=:=}, e.g.,
the equational constraint \ccode{head [] =:= head []}
is not satisfiable.\footnote{From now on, we use the
standard notation for lists, i.e., \code{[]} denotes the empty list
and \code{(x:xs)} denotes a list with head element \code{x}
and tail \code{xs}.}

Due to this constructive definition, \ccode{=:=}
can be considered as a binary function defined by the
following rules (here we show only the rules for the Boolean
and list types, where \code{Success} denotes the only constructor
of the type \code{Success} of constraints):
\begin{curry}
  True   =:= True    =  Success
  False  =:= False   =  Success
  []     =:= []      =  Success
  (x:xs) =:= (y:ys)  =  x =:= y & xs =:= ys

  Success & c  =  c
\end{curry}
%
If we translate these operations into Haskell by the scheme
shown in Section~\ref{sec:Compilation},
the following rules are added to these rules
in order to propagate choices and failures:
\begin{haskell}
  Fail         =:= _             =  Fail
  _            =:= Fail          =  Fail
  Choice i l r =:= y             =  Choice i (l =:= y) (r =:= y)
  x            =:= Choice i l r  =  Choice i (x =:= l) (x =:= r)
  _            =:= _             =  Fail

  Fail         & _     =  Fail
  _            & Fail  =  Fail
  Choice i l r & c     =  Choice i (l & c) (r & c)
  _            & _     =  Fail

\end{haskell}
%
Although this is a correct implementation of equational constraints,
it might lead to an unnecessarily large search space
when it is applied to generators representing logic variables.
For instance, consider the following generator for Boolean lists:
\begin{curry}
  aBoolList = [] ? (aBool : aBoolList)
\end{curry}
This is translated into Haskell as follows:
\begin{haskell}
  aBoolList :: IDSupply -> [Bool]
  aBoolList s = Choice (thisID s) [] (aBool (leftSupply s)
                                      : aBoolList (rightSupply s))
\end{haskell}
Now consider the equational constraint \ccode{x =:= [True]}.
If the logic variable \code{x} is replaced by \code{aBoolList},
the translated expression \ccode{aBoolList s =:= [True]}
creates a search space when evaluating its first argument,
although there is no search required since there is only one possible
binding for \code{x}.
Furthermore and even worse, there is an infinite search space
when unifying two logic variables. For instance,
the expression \ccode{xs =:= ys \& xs++ys =:= [True]}
results in an infinite search space when the logic variables
\code{xs} and \code{ys} are replaced by generators.

To avoid these problems, we have to implement the idea
of the well-known unification principle \cite{Robinson65}.
Instead of enumerating all values for logic variables
occurring in an equational constraint,
we \emph{bind} them to another variable or term.
Since we compile into a purely functional language,
the binding cannot be performed by some side effect.
Instead, we add binding constraints to the computed results.

To implement unification, we have to distinguish free variables
from ``standard choices'' (introduced by the operation \ccode{?})
in the target code. For this purpose, we refine the definition
of the type \code{ID} as follows:\footnote{For the sake
of simplicity, we consider in the following the
implementation of \code{IDSupply} as unbounded integers.}
\begin{haskell}
  data ID = ChoiceID Integer | FreeID Integer
\end{haskell}
The new constructor \code{FreeID} marks a choice corresponding to
a free variable, e.g., the generator for Boolean variables is
redefined as
\begin{haskell}
  aBool s = Choice (FreeID (thisID s)) True False
\end{haskell}
%
If an operation is applied to a free variable and requires its value,
the free variable is transformed into a standard choice.
For this purpose, we define a simple operation to perform
this transformation:
\begin{haskell}
  narrow :: ID -> ID
  narrow (FreeID i) = ChoiceID i
  narrow x          = x
\end{haskell}
Furthermore, we use this operation in narrowing steps,
i.e., in all rules operating on \code{Choice} constructors.
For instance, in the implementation of the operation \code{not}
we replace the rule
\begin{haskell}
  not (Choice i x1 x2) s = Choice i (not x1 s) (not x2 s)
\end{haskell}
by the rule
\begin{haskell}
  not (Choice i x1 x2) s = Choice (narrow i) (not x1 s) (not x2 s)
\end{haskell}
With this small change, the run-time system can distinguish
free variables (choices with an identifier of the form \code{(FreeID $\ldots$)})
from other values. This is quite useful at the top-level
to print values containing logic variables in a specific form
rather than enumerating all values, similarly to logic programming systems.
For instance, KiCS2 evaluates the application of \code{head}
to an unknown list as follows:
\begin{haskell}
  Prelude> head xs where xs free
  {xs = (_x2:_x3)} _x2
\end{haskell}
Here, free variables are marked by the prefix \code{\us{}x}.

As mentioned above, the consideration of free variables
is relevant in equational constraints where \emph{binding constraints} are
generated. For this purpose, we introduce a type to represent
a binding constraint as a pair of a choice identifier
and a decision for this choice:
\begin{haskell}
  data Constraint = ID :=: Decision
\end{haskell}
Furthermore, we extend each data type by the possibility to
add such a constraint:
\begin{haskell}
  data Bool   = $\ldots$ | Guard Constraint Bool
  data List a = $\ldots$ | Guard Constraint (List a)
\end{haskell}
Thus, \code{(Guard $c$ $v$)} represents a \emph{constrained value},
i.e., the value $v$ is only valid if the constraint $c$ is
consistent with the decisions previously made during search.
The binding constraints are created by the
equational constraint operation \ccode{=:=}: if a free variable
should be bound to a constructor, we make the same
decision as it would be done in the successful branch
of the generator. In case of Boolean values,
this can be implemented by the following additional rules
for \ccode{=:=}:
\begin{haskell}
  Choice (FreeID i) _ _ =:= True   =  Guard (i :=: ChooseLeft ) Success
  Choice (FreeID i) _ _ =:= False  =  Guard (i :=: ChooseRight) Success
\end{haskell}
Hence, the binding of a variable to some known value
is implemented as a binding constraint for the choice identifier
for this variable. However, if we want to bind a variable
to another variable, we cannot store a concrete decision.
Instead, we store the information that the decisions for
both variables, when they are made to extract values,
must be identical. For this purpose, we extend the \code{Decision}
type to cover this information:
\begin{haskell}
  data Decision = $\ldots$ | BindTo ID
\end{haskell}
Furthermore, we add the rule that an equational constraint
between two variables yields a binding for these variables:
\begin{haskell}
  Choice (FreeID i) _ _ =:= Choice (FreeID j) _ _
    =  Guard (i :=: BindTo j) Success
\end{haskell}
%
The consistency of constraints is checked when values are extracted
from a choice structure, e.g., by the operation \code{printValsDFS}.
For this purpose, we extend the definition of the corresponding search
operations by calling a constraint solver for the constraints.
For instance, the definition of \code{printValsDFS} is extended by
a rule handling constrained values:
%
\begin{haskell}
  $\ldots$
  printValsDFS (Guard c x) = do consistent <- add c
                                if consistent then do printValsDFS x
                                                      remove c
                                              else return ()
  $\ldots$
\end{haskell}
The operation \code{add} checks the consistency of the constraint \code{c}
with the decisions made so far and, in case of consistency,
stores the decision made by the constraint.
In this case, the constrained value is evaluated
before the constraint is removed to allow backtracking.
Furthermore, the operations \code{lookupDecision} and \code{setDecision}
are extended to deal with bindings between two variables,
i.e., they follow variable chains in case of \code{BindTo} constructors.

Due to the extensions described in this section,
we obtain an implementation of a unification procedure
that integrates the demand-driven evaluation of operations
occurring in the arguments.
Benchmarks evaluating the efficiency of this approach are
shown in Section~\ref{sec:Benchmarks}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Functional Patterns}
\label{sec:FuncPatterns}

A well-known disadvantage of equational constraints
is the fact that \ccode{=:=} is interpreted as strict equality.
Thus, if one uses equational constraints to express
requirements on arguments, the resulting operations might be too strict.
For instance, the equational constraint in the condition
defining \code{last} (see Section~\ref{ex:last})
requires that \code{ys++[e]} as well as \code{xs}
must be reducible to identical values so that the input list \code{xs}
is completely evaluated.
Hence, if \code{failed} denotes an operation whose evaluation fails,
the evaluation of \code{last [failed,True]} has no result.
On the other hand, the evaluation of \code{last' [failed,True]} yields
the value \code{True}, i.e., the definition of \code{last'} is less strict.
This is due to the fact that a functional pattern like
\code{(xs++[e])} abbreviates all values to which it can be evaluated
(by narrowing), like \code{[e]}, \code{[x1,e]}, \code{[x1,x2,e]} etc.
Conceptually, the rule defining \code{last'}
abbreviates the following (infinite) set of rules:
\begin{curry}
  last' [e] = e
  last' [x1,e] = e
  last' [x1,x2,e] = e
  $\ldots$
\end{curry}
As a further example,
consider an operation that returns the first duplicated element in
a list.
Using equational constraints, we can define it as follows:
\begin{curry}
  fstDup xs | xs =:= ys++[e]++zs & elem e ys =:= True & nub ys =:= ys
            = e    where ys, zs, e free
\end{curry}
The first equational constraint is used to split the input list \code{xs}
into three sublists.
The last constraint ensures that the first sublist
\code{ys} does not contain duplicated elements
(the library operation \code{nub} removes all duplicates from a list)
and the second constraint ensures that the first element after \code{ys}
occurs in \code{ys}.
Although this implementation is concise, it cannot be applied
to infinite lists due to the strict interpretation of \ccode{=:=}.
This is not the case if we define this operation by a functional pattern:
\begin{curry}
  fstDup' (ys++[e]++zs) | elem e ys =:= True & nub ys =:= ys
                        = e
\end{curry}
%
Functional patterns are quite useful in situations
where one wants to select parts of some structure
at arbitrary positions. For instance,
functional patterns have been applied to support
the high-level programming of querying or transforming
XML data \cite{Hanus11ICLP}.

Obviously, one cannot implement functional patterns by
a transformation into an infinite set of rules. Instead, they are
implemented by a specific \emph{lazy unification} procedure \ccode{=:<=}
\cite{AntoyHanus05LOPSTR}.
For instance, the definition of \code{last'} is transformed into
\begin{curry}
  last' ys | (xs++[e]) =:<= ys  = e   where xs, e free
\end{curry}
The behavior of \ccode{=:<=} is similar to \ccode{=:=}
except for the case that a variable in the left argument
should be bound to some expression: instead of evaluating
the expression to some value and binding the variable to the value,
the variable is bound to the \emph{unevaluated} expression
(see \cite{AntoyHanus05LOPSTR} for more details).
Due to this slight change, failures or infinite structures
in actual arguments do not cause problems in the matching
of functional patterns.

The following discussion shows the requirements on an 
implementation of functional patterns.
The general structure
is quite similar to equational constraints
with the exception that variables could be also bound
to unevaluated expressions. Only if such variables are later
accessed, the expressions they are bound to are evaluated.
This can be achieved by adding a further alternative
to the type of decisions:
\begin{haskell}
  data Decision = $\ldots$ | LazyBind Constraint
\end{haskell}
%
The implementation of the lazy unification operation \ccode{=:<=}
is almost identical to the strict unification operation \ccode{=:=}
as shown in Section~\ref{sec:Unification}.
The only difference is in the rules where a free variable occurs
in the left argument. All these rules are replaced by the single
rule
\begin{haskell}
  Choice (FreeID i) _ _ =:<= x
    = Guard (i :=: LazyBind (lazyBind i x)) Success
\end{haskell}
where the auxiliary operation \code{lazyBind}
implements the demand-driven evaluation of the right argument \code{x}:
\begin{haskell}
  lazyBind :: ID -> a -> Constraint
  lazyBind i True  = i :=: ChooseLeft
  lazyBind i False = i :=: ChooseRight
\end{haskell}
The use of the additional \code{LazyBind} constructor allows the argument 
\code{x} to be stored in a binding constraint without evaluation 
(due to the lazy evaluation strategy of the
target language Haskell).
However, it is evaluated by \code{lazyBind} when its binding
is required by another part of the computation.
Similarly to equational constraints,
lazy bindings are processed by a solver when values are extracted.
In particular, if a variable has more than one lazy binding constraint
(which is possible if a functional pattern evaluates to a non-linear term),
the corresponding expressions are evaluated and unified
according to the semantics of functional patterns
\cite{AntoyHanus05LOPSTR}.

In order to demonstrate the operational behavior of our implementation,
we sketch the evaluation of the lazy unification constraint
\code{xs++[e] =:<= [failed,True]} that occurs when the expression
\code{last' [failed,True]} is evaluated (here we omit failed branches
and some other details; note that logic variables are replaced
by generators):
\begin{haskell}
  aBoolList 2 ++ [aBool 3] =:<= [failed,True]
  $\leadsto$ [aBool 4,aBool 3] =:<= [failed,True]
  $\leadsto$ aBool 4 =:<= failed & aBool 3 =:<= True & [] =:<= []
  $\leadsto$ Guard (4 :=: LazyBind (lazyBind 4 failed))
           (Guard (3 :=: LazyBind (lazyBind 3 True)) Success)
\end{haskell}
If the binding of identifier \code{3} is later required,
\code{(lazyBind 3 True)} is evaluated to \code{[3 :=: ChooseLeft]}
which corresponds to the value \code{True} of the generator \code{(aBool 3)}.
Note that the variable with identifier \code{4} does not occur
anywhere else so that the binding \code{(lazyBind 4 failed)}
will never be evaluated, as intended.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Benchmarks}
\label{sec:Benchmarks}

% The benchmark results are from unif@climens2011_7_5_20_38_47.bench

In this section we evaluate our implementation by some benchmarks.
They were executed on a Linux machine running Debian 5.0.7 with
an Intel Core 2 Duo (3.0GHz) processor.
KiCS2 has been executed with
the Glasgow Haskell Compiler (GHC 7.0.4, option -O2).
For a comparison with other mature implementations of Curry,
we considered PAKCS \cite{Hanus10PAKCS}
(Version 1.9.2, based on a SICStus-Prolog 4.1.2)
and MCC \cite{Lux99FLOPS} (Version 0.9.10). The timings were performed with the
time command measuring the execution time (in seconds) of a compiled
executable for each benchmark as a mean of three runs.
The programs used for the benchmarks,
partially taken from \cite{AntoyHanus05LOPSTR},
are \code{last} (compute the last element of a list),\footnote{%
\ccode{inc x n} is a naive addition that \code{n} times
increases its argument \code{x} by 1.}
\code{simplify} (simplify a symbolic arithmetic expression),
\code{varInExp} (non-deterministically return a variable occuring
                  in a symbolic arithmetic expression),
\code{half} (compute the half of a peano number using logic variables),
and
\code{palindrome} (check whether a list is a palindrome).

\begin{figure}[t]
\centering
\begin{tabular}{|l|r|r|r|}
\hline
Expression                              & \code{==~}& \code{=:=} & \code{=:<=} \\
\hline
\code{last (map (inc 0) [1..100000])}   & 134.23    & 0.83      & 0.26        \\
% \code{grep}                             & 1.06      & 0.19      &             \\
\code{simplify}                         & 8.50      & 7.15      & 7.34        \\
\code{varInExp}                         & 11.72     & 1.50      & 0.14        \\
% \code{varInExp2}                        & 423.01    & 10.50     & 1.04        \\
% \code{varInExp3}                        & 399.68    & 16.70     & 11.28       \\
\code{fromPeano (half (toPeano 10000))} & 32.92     & 6.35      & 11.88       \\
\code{palindrome}                       & 28.59     & 15.58     & 23.21       \\
% \code{horseman}                         & 3.58      & 3.13      &             \\
\hline
\end{tabular}
\caption{Benchmarks: comparing different equation solving methods}
 \label{fig:search-strategies}
\end{figure}

In Section~\ref{sec:Unification} we mentioned that equational
constraints could also be solved by generators without variable bindings,
but this technique might increase the search space due to the
possibly superfluous generation of all values.
To show the beneficial effects of our implementation
of equational constraints with variable bindings,
we compare in Fig.~\ref{fig:search-strategies}
the results of using equational constraints (\code{=:=})
to the results
where the Boolean equality operator (\code{==}) is used (which
does not perform bindings but enumerate all values).
As expected, the creation and traversal of a large search space introduced
by \code{==} is in most cases much slower than our presented approach
with variable bindings.
In addition, the example \code{last} shows that the lazy 
unification operator (\code{=:<=}) improves performance when unifying an
expression which has to be evaluted only partially. Using strict unification,
all elements of the list are (unnecessarily) evaluated.

\begin{figure}[t]
\centering
\begin{tabular}{|l|r|r|r|}
\hline
Expression                              & KiCS2 & PAKCS & MCC  \\
\hline
\code{last (map (inc 0) [1..100000])}   & 0.83  & 4.25   & 0.14 \\
% \code{grep}                             & 0.19  & 2.88   & 0.14 \\
\code{simplify}                         & 7.15  & 0.15   & 0.00 \\
\code{varInExp}                         & 1.50  & 3.89   & 0.30 \\
% \code{varInExp2}                        & 10.50 & 3.48   & stckOflw \\
% \code{varInExp3}                        & 16.70 & 19.22  & 1.18 \\
\code{fromPeano (half (toPeano 10000))} & 6.35  & 108.88 & 3.22 \\
\code{palindrome}                       & 15.58 & 32.56  & 1.07 \\
% \code{horseman}                         & 3.13  & 8.70   & 0.42 \\
\hline
\end{tabular}
\caption{Benchmarks: equational constraints in different Curry implementations}
 \label{fig:unification}
\end{figure}

In contrast to the Curry implementations PAKCS and MCC, our approach is based
on an explicit representation of the search space instead of
backtracking and manipulating
a global state containing bindings for logic variables.
Nevertheless, the benchmarks in Fig.~\ref{fig:unification}
show that it can compete with or even outperform these implementations.
In these benchmarks only equational constraints are used.
The results show that the implementation of unification
of MCC performs best. However, in most cases
our implementation outperforms the Prolog-based PAKCS implementation,
except for some examples where further investigations
and optimizations are necessary.

\begin{figure}[t]
\centering
\begin{tabular}{|l|r|r|}
\hline
Expression                              & KiCS2 & PAKCS    \\
\hline
\code{last (map (inc 0) [1..100000])}   & 0.26  & 3.97     \\
\code{simplify}                         & 7.34  & 0.27     \\
\code{varInExp}                         & 1.40  & 8.88     \\
% \code{varInExp2}                         & 1.04  & 0.08    \\
% \code{varInExp3}                         & 11.28 & 16.62   \\
\code{fromPeano (half (toPeano 10000))} & 11.88 & $\infty$ \\
\code{palindrome}                       & 23.21 & $\infty$ \\
\hline
\end{tabular}
\caption{Benchmarks: comparing different implementations of functional patterns}
 \label{fig:functional-patterns}
\end{figure}

As MCC does not support functional patterns, the performance of lazy
unification is compared with PAKCS only (Fig.~\ref{fig:functional-patterns}).
Again, our compiler performs well against PAKCS
and outperforms it in most cases (``$\infty$'' denotes a run time
of more than 30 minutes). \todo{extend}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions and Related Work}
\label{sec:Conclusions}

We have presented an implementation of equational constraints
and functional patterns in a purely functional implementation
of Curry. Our implementation is based on adding binding constraints
to computed values and processing them when values are extracted
at the top level of a computation.
Since we have added only new constructors and rules
for them in our implementation, it does not introduce
additional work if these features are not used, i.e.,
our implementation does not sacrifice the high efficiency
of the kernel implementation as shown in
\cite{BrasselFischerHanusReck11}.
However, if these features are used, they usually lead
to a more efficient execution, as demonstrated by our benchmarks.

Other implementations of equational constraints in functional logic
languages use side effects for their implementation.
For instance, PAKCS \cite{Hanus10PAKCS}
exploits the implementation of logic variables in Prolog,
which are implemented on the primitive level by side effects.
MCC \cite{Lux99FLOPS} compiles into C where a specific
abstract machine implements the handling of logic variables.
We have shown that our implementation is competitive to those.
In contrast to those systems, our implementation supports
a variety of search strategies, like breadth-first or parallel search,
where the avoidance of side effects is important.

For future work it might be interesting to add
further constraint structures, like real arithmetic
or finite domain constraints, to our implementation.
This might be possible by extending the kinds of constraints
of our implementation and solving them by
functional approaches like \cite{SchrijversStuckeyWadler09}.

\bibliographystyle{mhnumber} %plain}
\bibliography{mh}

\end{document}

% LocalWords:  Curry KiCS equational