### 2012-06-08 talks/badhonnef2012/set-functions/talk.tex

\documentclass[utf8]{beamer}

\usepackage{graphicx}
\usepackage{color}
\usepackage{listings}
\usetheme{Boadilla}

\lstset{basicstyle=\small}

\setbeamertemplate{footline}%{infolines theme}
{
\leavevmode%
\hbox{%
\begin{beamercolorbox}[wd=.333333\paperwidth,ht=2.25ex,dp=1ex,center]{author in head/foot}%
\usebeamerfont{author in head/foot}\insertshortauthor%~~(\insertshortinstitute)
\end{beamercolorbox}%
\begin{beamercolorbox}[wd=.333333\paperwidth,ht=2.25ex,dp=1ex,center]{title in head/foot}%
\usebeamerfont{title in head/foot}\insertshorttitle
\end{beamercolorbox}%
\begin{beamercolorbox}[wd=.333333\paperwidth,ht=2.25ex,dp=1ex,right]{date in head/foot}%
\usebeamerfont{date in head/foot}\insertshortdate{}\hspace*{2em}
\insertframenumber{} / \inserttotalframenumber\hspace*{2ex}
\end{beamercolorbox}}%
\vskip0pt%
}

\begin{document}

\title{Negation as Failure with Set Functions}

\date{May 04, 2012}

\author[M. Hanus, B. Peemöller, \underline{F. Reck}]
 { Michael Hanus \and Björn Peemöller\\
            \underline{Fabian Reck} \\
  \texttt{\{mh, bjp, fre\}@informatik.uni-kiel.de}}

\institute{Kiel University}

\begin{frame}
\titlepage
\end{frame}

\section{Introduction}

\begin{frame}[fragile]
\frametitle{Curry}
\begin{columns}[t]
\column{.45\textwidth}
\begin{itemize}
\item Lazy functional logic programing language.
\item Haskell-like syntax.
\item Extended by non-determinism, free variables, constraints, unification,
      failure.
\end{itemize}
\begin{example}<2>
\begin{lstlisting}
> ensure aBool
True
\end{lstlisting}
\end{example}
\column{.45\textwidth}
\begin{block}{Running Code Example}
\begin{lstlisting}
data Bool = True | False

aBool :: Bool
aBool = True ? False

not True  = False
not False = True

ensure True = True
\end{lstlisting}
\end{block}
\end{columns}
\end{frame}

\begin{frame}[fragile]
\frametitle{Example}
\begin{block}{Flight data as non-deterministic operation}
\begin{lstlisting}[mathescape]
flight = (LH469, Portland, Frankfurt, 10:.15)
       ? (NWA92, Portland, Amsterdam, 10:.00)
       ? (LH10,  Frankfurt,Hamburg,    1:.00)
       ? (KL1783,Amsterdam,Hamburg,    1:.52)

itinerary orig dest 
   | flight =:= (num,orig,dest,len)
   = [num]                         where num,len free

itinerary orig dest 
   | flight =:= (num1,orig,stop,len1)
   & flight =:= (num2,stop,dest,len2)
   = [num1,num2]      where num1,len1,num2,len2,stop free
\end{lstlisting}
\end{block}
\pause
Problem: itinerary with shortest air time? $\leadsto$ reason about \emph{all} values
\end{frame}

\begin{frame}[fragile]
\frametitle{Encapsulated Search}
\begin{itemize}
\item Allows to access the non-deterministic results of an expression
      within the program
\end{itemize}
\begin{block}{Example}
\begin{lstlisting}
> allValues aBool
[True,False]
\end{lstlisting}
\end{block}
\pause
\begin{block}{Problem}
\begin{lstlisting}
> allValues $! aBool
[True]
[False]
\end{lstlisting}%$
\end{block}
\begin{itemize}
\item Result depends on the evaluation order
\item Declarative languages are order independent 
\end{itemize}
\end{frame}

\begin{frame}[fragile]
\frametitle{Set Functions}
\begin{itemize}
\item Novel approach to encapsulated search
\item For every function $f$ there is a function $f_S$
\item $f_S$ takes the same arguments as $f$
\item The result of $f_S$: Set of non-deterministic results of $f$
\item Non-determinism in arguments to $f_S$ is not capsuled
\item Every result of $f \overline{t}$ is element of a
      result of $f_S \overline{t}$
\end{itemize}
\begin{block}{Example}
\begin{lstlisting}[mathescape]
> aBool$_S$ 
[True,False]

> not$_S$ aBool
[False]
[True]
\end{lstlisting}
\end{block} 
\end{frame}

\begin{frame}[fragile]
\frametitle{Implementation Approaches}
\begin{itemize}
\item Evaluate the arguments strictly \emph{before}
      passing them to the function
\pause
\begin{block}{Problem}
\begin{lstlisting}[mathescape]
answer x = 42

> answer (ensure False)
42

> answer$_S$ (ensure False)
no more solutions
\end{lstlisting}
\end{block}
\item Identify non-determinism that stems from an argument
\end{itemize}
\end{frame}

\section{Compilation scheme}

\begin{frame}[fragile]
\frametitle{Representation of Non-Determinism and Failure in KiCS2}
\begin{itemize}
  \item No built-in non-determinism in Haskell.
  \item Idea: Explicit representation in data types as constructors.
\end{itemize}
\pause
\begin{block}{Curry}
\begin{lstlisting}
data Bool = True | False
aBool = True ? False
ensure True = True
\end{lstlisting}
\end{block}

\begin{block}{Haskell}
\begin{lstlisting}
data Bool = True | False | Choice Bool Bool | Fail
aBool = Choice True False
ensure True = True
ensure (Choice l r) = Choice (ensure l) (ensure r)
ensure _    = Fail
\end{lstlisting}
\end{block}
\end{frame}

\begin{frame}[fragile]
\frametitle{Implementation of Set Functions}
\begin{itemize}
\item Mark choices that stem from arguments
\begin{block}{Curry}
\begin{lstlisting}[mathescape]
answer$_S$ x = search (answer (cover x))

cover True         = True
cover False        = False
cover (Choice l r) = Choice$^C$ (cover l) (cover r)
cover Fail         = Fail

search (Choice l r)  = search l ++ search r
search (Choice$^C$ l r) = Choice (search l) (search r)
search Fail          = []
search x             = [x]
\end{lstlisting}
\end{block}
\item Marked choices are retained 
\end{itemize}
\end{frame}

\begin{frame}[fragile]
\frametitle{Negation as Failure}
\begin{itemize}
\item Just check if the result of a set function is the empty set
\begin{block}{Example}
\begin{lstlisting}[mathescape]
data Size = Small | Big
aSize = Small ? Big

getSmallest | isEmpty (smallerThan$_S$ size) = size
  where size = aSize

smallerThan size | anotherSize < size = anotherSize
  where anotherSize = aSize

> getSmallest
Small
\end{lstlisting}
\end{block}
\end{itemize}
\end{frame}

\begin{frame}[fragile]
\frametitle{Negation as Failure and Failures}
\begin{block}{Problem}
\begin{lstlisting}[mathescape]
getSmallest | isEmpty (smallerThan$_S$ 
                         (size =:= aSize &> size)) 
            = size
  where size = aSize

> getSmallest
Small
Small
Big
\end{lstlisting}
\end{block}
\begin{itemize}
 \item Failures in arguments lead to empty results of set functions
 \item Set functions should behave as if arguments were
       evaluated outside
 \item Implementation: Also mark failures
\end{itemize}
\end{frame}

\begin{frame}[fragile]
\frametitle{Be Lazy}
\begin{block}{Problem}
\begin{lstlisting}[mathescape]
trueOrSomething x = True ? x
> trueOrSomething failed
True
> trueOrSomething$_S$ failed
no more solutions
\end{lstlisting}
\end{block}
\begin{itemize}
\item Set functions should behave as if arguments were
      evaluated ouside as far as all their non-deterministic
      branches require
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Conclusion and future work}
\begin{itemize}
\item Set functions are an order independent approach to encapsulated
      search
\item Check for empty result set allows programming with negation as failure
\item Difficult to handle failing arguments
\item Need for theoretic justification of the behaviour
\item Need for formal semantics
\end{itemize}
\end{frame}

\end{document}