### 2014-10-06 benchmarks/benchmark_papers/kics2/bench_kics2.tex
Benchmark suite updated and benchmark paper added

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% This is a paper to benchmark various versions of the KiCS2 system.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass{article}

\usepackage{currycode}
\usepackage{graphicx}

\begin{document}
\sloppy

\title{Benchmarking KiCS2}
\author{Michael Hanus}
\maketitle

\section{Introduction}

This document contains the result of benchmarking various
Curry systems.
%
All benchmarks were executed in the following environment:
\begin{itemize}
\item \runcurry{getOS} machine (named ``\runcurry{getHostName}'')
\item operating system: \runcurry{getSystemID} \runcurry{getSystemRelease}
\item CPU: ``\runcurry{getCPUModel}'' processor containing
      \runcurry{getCoreNumber} cores
\item time limit for each benchmark: \runcurry{return (show timeLimit)} seconds
\item number of iterations for each benchmark:
       \runcurry{return (show numberOfRuns)}
\end{itemize}

\section{Benchmarks}

%%% The actual benchmark programs are in a separate Curry module:
\begin{curryprog}
----------------------------------------------------------------------
import Benchmarks
import BenchKiCS2
----------------------------------------------------------------------
\end{curryprog}

\subsection{Deterministic Benchmarks}

The following diagram compares MCC, PAKCS, and KiCS2
on the classical naive reverse benchmark for different list lengths:

\begin{center}
\runcurry{benchNRevAndPlot 5.2 [500,1000..] "plot_all"}
\end{center}

Next we test various Curry systems on deterministic operations,
i.e., operations without overlapping rules or free variables:

\begin{center}
%\runcurry{mainBench (MCC:PAKCS:kics2Systems) detBench}
\end{center}

\subsection{Non-deterministic DFS Benchmarks}

The following benchmark programs contain non-deterministic operations
or free variables. We evaluate the computation of a
first solution/value via DFS in PAKCS vs. KiCS2:

\begin{center}
%\runcurry{mainBench [PAKCS, KiCS2 ""] nondetBenchDFS}
\end{center}

\subsection{Non-deterministic Benchmarks}

\begin{center}
%\runcurry{mainBench kics2Systems nondetBench}
\end{center}

\subsection{Encapsulated Search Benchmarks}

\begin{center}
%\runcurry{mainBench kics2Systems encapsBench}
\end{center}

\end{document}