### 2015-11-17 docs/src/manual.tex
Manual updated, documentation of tools moved into currytools repo

\documentclass[11pt,fleqn]{article}

\usepackage{latexsym}
\usepackage{makeidx}
\usepackage{url}
\usepackage{xspace}
\usepackage{graphicx}

\input{version}

% define home dir of the Curry system in order to locate tool manuals:
\newcommand{\curryhome}{../..}

%%% ------------------------------------------------------------------

\usepackage[colorlinks,linkcolor=blue]{hyperref}
\hypersetup{bookmarksopen=true}
\hypersetup{bookmarksopenlevel=0}
\hypersetup{pdftitle={KiCS2: The Kiel Curry System Version 2}}
\hypersetup{pdfauthor={Michael Hanus}}
%\hypersetup{pdfstartview=Title}
\hypersetup{pdfstartview=FitH}
\usepackage{thumbpdf}

%%% ------------------------------------------------------------------

\setlength{\textwidth}{16.5cm}
\setlength{\textheight}{23cm}
\renewcommand{\baselinestretch}{1.1}
\setlength{\topmargin}{-1cm}
\setlength{\oddsidemargin}{0cm}
\setlength{\evensidemargin}{0cm}
\setlength{\marginparwidth}{0.0cm}
\setlength{\marginparsep}{0.0cm}

\newlength{\figurewidth}
\setlength{\figurewidth}{\textwidth}
\addtolength{\figurewidth}{-0.4cm}

% font for program texts
\renewcommand{\tt}{\usefont{OT1}{cmtt}{m}{n}\selectfont}
\newcommand{\codefont}{\small\tt}

\usepackage{listings}
\lstset{aboveskip=1.5ex,
        belowskip=1.5ex,
        showstringspaces=false, % no special string space
        mathescape=true,
        flexiblecolumns=false,
        xleftmargin=2ex,
        basewidth=0.52em,
        basicstyle=\small\ttfamily}

\lstset{literate={->}{{$\rightarrow{}\!\!\!$}}3
       }

\lstnewenvironment{curry}{}{}
\lstnewenvironment{currynomath}{\lstset{mathescape=false}}{} % Curry w/o math
\newcommand{\listline}{\vrule width0pt depth1.75ex}

% program text in normal text
\newcommand{\code}[1]{\mbox{\codefont #1}}

% program text in normal text with apostrophs
\newcommand{\ccode}[1]{``\code{#1}''}

\newcommand{\pindex}[1]{\index{#1@{\tt #1}}}  % program elements in index

\newcommand{\todo}[1]{\fbox{\sc To do: #1}}
\newcommand{\CYS}{KiCS2\xspace} % name of Curry system described here
\newcommand{\cyshome}{\mbox{\textit{kics2home}}\xspace} % symbolic installation directory
\newcommand{\cb}{CurryBrowser\xspace}
\newcommand{\curryrc}{.kics2rc} % name of the system's rc file

\usepackage{syntax}

% allow underscores in programs:
\catcode`\_=\active
\let_=\sb
\catcode`\_=12

\newcommand{\translation}[1]{\bigskip\fbox{\begin{minipage}{15cm}\textbf{Translation: }{#1}\end{minipage}}\par\bigskip}

% produce an index:
\makeindex

\begin{document}
\sloppy

\begin{titlepage}
\pdfbookmark[1]{Title}{Title}
\begin{center}
\fbox{
\begin{minipage}[t]{\figurewidth}
\begin{center}\vspace{10ex}
{\Huge\bf KiCS2}\\[4ex]
{\huge The Kiel Curry System (Version 2)}\\[7ex]
{\huge User Manual}\\[4ex]
\kicsversiondate\\[6ex]
\Large
Michael Hanus$^1$ [editor] \\[3ex]
{\large Additional Contributors:}\\[2ex]
Bernd Bra\ss{}el$^2$ \\
Bj\"orn Peem\"oller$^3$\\
Fabian Reck$^4$ \\
Jan Rasmus Tikovsky$^5$ \\[4ex]
\normalsize
(1) University of Kiel, Germany, {\tt mh@informatik.uni-kiel.de} \\
(2) University of Kiel, Germany, {\tt bbr@informatik.uni-kiel.de} \\
(3) University of Kiel, Germany, {\tt bjp@informatik.uni-kiel.de} \\
(4) University of Kiel, Germany, {\tt fre@informatik.uni-kiel.de} \\
(4) University of Kiel, Germany, {\tt jrt@informatik.uni-kiel.de} \\[5ex]~
\end{center}
\end{minipage}}
\end{center}
\end{titlepage}

\pdfbookmark[1]{Contents}{Contents}
\tableofcontents

\include{preface}
\include{overview}
\include{usage}
\include{extensions}
\include{syntax}
\clearpage
\input{\curryhome/currytools/currydoc/Docs/manual}
\clearpage
\input{\curryhome/currytools/browser/Docs/manual}
\clearpage
\input{\curryhome/currytools/currytest/Docs/manual}
\clearpage
\input{\curryhome/currytools/erd2curry/Docs/manual}
\clearpage
\input{\curryhome/currytools/spicey/Docs/manual}
\include{problems}

% Bibliography
\addcontentsline{toc}{section}{Bibliography}
\bibliography{mh}
\bibliographystyle{plain}

\appendix

\include{libraries}
\include{markdown_syntax}
\include{auxiliary_files}
\include{external_ops}

% Index
\addcontentsline{toc}{section}{Index}
\printindex

\end{document}

% LocalWords:  KiCS